Manuscript_Github_Pchabaudi_compBerghei

### 2022-08-09 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_compBerghei.Rmd
Changed figures to not show ancestral as comparing S vs L proportions

```{r chab_subfam_normByGeneNum}
chab_tpm_subfam.df.long <- chab_tpm.df %>% 
  rownames_to_column('Geneid') %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(subfam = cir_info$subfam_num[match(Geneid, cir_info$Geneid)],
         transmission = str_extract(sample, pattern = 'rMT|SBP'),
         time = str_extract(sample, '[[:digit:]]{2}h|sch.culture'),
         name = paste0(time, str_extract(sample, 
                                pattern = '(_[[:digit:]]$)')))

ggplot(chab_tpm_subfam.df.long, 
       aes(x = factor(name,
                      levels = ordering(name, 
                                        order_vector = c(times_stages_table$time, 'sch.culture'))),
           y = tpm,
           fill = subfam)) +
  geom_col(position = "fill") +
  scale_fill_manual(values = clade_col_num) +
  facet_wrap(~transmission, scales = "free_x")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))
```

@@

```{r berghei_subfam}
berg_tpm_subfam.df.long <- berghei_tpm[,
                                       c('Geneid',
                                         str_subset(colnames(berghei_tpm), 
                                                    pattern = 'Asex'))] %>% 
  filter(Geneid %in% bir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(subfam = bir_clade$subfam_num[match(Geneid, bir_clade$Geneid)],
         experiment = str_extract(sample, pattern = '^E[[:digit:]]{2}'),
         stage = str_extract(sample, 'Asex.+(?=_[[:digit:]])'),
         name = str_extract(sample, '(?<=Asex\\.).+'))

ggplot(berg_tpm_subfam.df.long, 
       aes(x = factor(name,
                      levels = ordering(name,
                                        order_vector = c('Ring','Troph','Schiz','Mixed'))),
           y = tpm,
           fill = subfam)) +
  geom_col(position = "fill") +
  xlab('stages') +
  scale_fill_manual(values = bir_clade_col_num) +
  facet_wrap(~experiment, scales = "free_x")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))
```

@@

>>>>

```{r chab_subfam_normByGeneNum}
chab_tpm_subfam.df.long <- chab_tpm.df %>% 
  rownames_to_column('Geneid') %>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(subfam = cir_info$subfam_num[match(Geneid, cir_info$Geneid)],
         SorL = cir_info$SorL[match(Geneid, cir_info$Geneid)],
         transmission = str_extract(sample, pattern = 'rMT|SBP'),
         time = str_extract(sample, '[[:digit:]]{2}h|sch.culture'),
         name = paste0(time, str_extract(sample, 
                                pattern = '(_[[:digit:]]$)')))

ggplot(filter(chab_tpm_subfam.df.long, SorL != 'ancestral'), 
       aes(x = factor(name,
                      levels = ordering(name, 
                                        order_vector = c(times_stages_table$time, 'sch.culture'))),
           y = tpm,
           fill = subfam)) +
  geom_col(position = "fill") +
  scale_fill_manual(values = clade_col_num) +
  facet_grid(~transmission, scales = "free_x") +
  # facet_grid(transmission ~ SorL, scales = "free")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 3))
```

@@

```{r berghei_subfam}
berg_tpm_subfam.df.long <- berghei_tpm[,
                                       c('Geneid',
                                         str_subset(colnames(berghei_tpm), 
                                                    pattern = 'Asex\\.(?!Mixed)'))] %>% 
  filter(Geneid %in% bir_id) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(subfam = bir_clade$subfam_num[match(Geneid, bir_clade$Geneid)],
         experiment = str_extract(sample, pattern = '^E[[:digit:]]{2}'),
         stage = str_extract(sample, 'Asex.+(?=_[[:digit:]])'),
         name = str_extract(sample, '(?<=Asex\\.).+'))

ggplot(filter(berg_tpm_subfam.df.long, subfam != 'ancestral (n = 1)'), 
       aes(x = factor(name,
                      levels = ordering(name,
                                        order_vector = c('Ring','Troph','Schiz','Mixed'))),
           y = tpm,
           fill = subfam)) +
  geom_col(position = "fill") +
  xlab('stages') +
  scale_fill_manual(values = bir_clade_col_num) +
  facet_wrap(~experiment, scales = "free_x")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))
```

@@

Hypothesis: There is a difference in the transcription level of L-form pirs as a proportion of all S/L-form pirs between Asexual Schizont samples and non-schizont samples

```{r chab_prop_test}
chab_tpm_subfam.L.df.long <- chab_tpm_subfam.df.long %>% 
  filter(!subfam == 'ancestral (n = 1)') %>% 
  mutate(subfam_L = grepl(subfam, pattern = '^L')) %>% 
  group_by(transmission, time, name, sample, subfam_L) %>% 
  summarise(sum_tpm = sum(tpm)) %>% 
  group_by(sample) %>% 
  mutate(sum_tpm_all = sum(sum_tpm)) %>% 
  filter(subfam_L)

chab_tpm_subfam.L.df.long %>% 
  group_by(sample) %>% 
  mutate(prop_L = sum_tpm/sum_tpm_all) %>% 
  arrange(desc(prop_L))

bind_rows(
  lapply(str_subset(unique(chab_tpm_subfam.L.df.long$sample), 
                    pattern = 'sch.culture',
                    negate = TRUE),
         function(sample.name){
           bind_rows(
             lapply(c('rMT_sch.culture_1', 
                      'rMT_sch.culture_2', 
                      'rMT_sch.culture_3'), 
                    function(sch.cult){
                      L_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm[
                        chab_tpm_subfam.L.df.long$sample == sample.name
                      ],
                      chab_tpm_subfam.L.df.long$sum_tpm[
                        chab_tpm_subfam.L.df.long$sample == sch.cult
                      ])
                      SorL_forms <- c(chab_tpm_subfam.L.df.long$sum_tpm_all[
                        chab_tpm_subfam.L.df.long$sample == sample.name
                      ],
                      chab_tpm_subfam.L.df.long$sum_tpm_all[
                        chab_tpm_subfam.L.df.long$sample == sch.cult
                      ])
                      test.stat <- prop.test(L_forms, SorL_forms,
                                             alternative = 'less')
                      data.frame(comp = paste0(sample.name, '-', sch.cult),
                                 pval = test.stat$p.value)
                    })
           )
         })
)%>% 
  mutate(padj = round(p.adjust(pval), digits = 5),
         pval = round(pval, digits = 5)) %>% 
  arrange(desc(padj))
```

```{r berg_prop_test}
berg_tpm_subfam.L.df.long <- berg_tpm_subfam.df.long %>% 
  filter(!subfam == 'ancestral (n = 1)') %>% 
  mutate(subfam_L = grepl(subfam, pattern = '^L')) %>% 
  group_by(experiment, name, sample, subfam_L) %>% 
  summarise(sum_tpm = sum(tpm)) %>% 
  group_by(sample) %>% 
  mutate(sum_tpm_all = sum(sum_tpm)) %>% 
  filter(subfam_L)

berg_tpm_subfam.L.df.long %>% 
  group_by(sample) %>% 
  mutate(prop_L = sum_tpm/sum_tpm_all) %>% 
  arrange(desc(prop_L))

bind_rows(
  lapply(str_subset(unique(berg_tpm_subfam.L.df.long$sample), 
                    pattern = 'Schiz|Mixed',
                    negate = TRUE),
         function(sample.name){
           bind_rows(
             lapply(str_subset(unique(berg_tpm_subfam.L.df.long$sample), 
                               pattern = 'Schiz'), 
                    function(sch.cult){
                      L_forms <- c(berg_tpm_subfam.L.df.long$sum_tpm[
                        berg_tpm_subfam.L.df.long$sample == sample.name
                      ],
                      berg_tpm_subfam.L.df.long$sum_tpm[
                        berg_tpm_subfam.L.df.long$sample == sch.cult
                      ])
                      SorL_forms <- c(berg_tpm_subfam.L.df.long$sum_tpm_all[
                        berg_tpm_subfam.L.df.long$sample == sample.name
                      ],
                      berg_tpm_subfam.L.df.long$sum_tpm_all[
                        berg_tpm_subfam.L.df.long$sample == sch.cult
                      ])
                      test.stat <- prop.test(L_forms, SorL_forms,
                                             alternative = 'less')
                      data.frame(comp = paste0(sample.name, '-', sch.cult),
                                 pval = test.stat$p.value)
                    })
           )
         })
)%>% 
  mutate(padj = round(p.adjust(pval), digits = 5),
         pval = round(pval, digits = 5)) %>% 
  arrange(desc(padj))
```
